On the presentation of orthogonal groups over polynomial rings

### 2019-06-08 dl-paper.tex
Reorganizing proof

We need to introduce notation for certain relations of $\St(\Phi, B)$. 
\begin{align}
&\,\,x_{\alpha}(a)x_{\alpha}(a') =  x_{\alpha}(a+a'),& \tag{R$1_d$} \\
&\,[x_{\alpha}(a),\,x_{\beta}(b)]= x_{\alpha+\beta}(N_{\alpha, \beta} \cdot ab),& \tag{R$2_{d,e}$} \\
&[x_{\alpha}(a),\,x_{\gamma}(b)]= 1,& \alpha \perp \beta,\ \tag{R$3^\bot_{d,e}$} \\
&[x_{\alpha}(a),\,x_{\beta}(b)]= 1,& \alpha - \beta \in \Phi,\ \tag{R$3^\angle_{d,e}$}
\end{align}
where $a, a' \in B_d$, $b\in R_e$. 
%For such a ring $R$ one can assume in the definition of the Steinberg group that the generators $x_\alpha(a)$ and relations R1--R3 are defined only for homogeneous elements.

For $n\geq 1$ we define ``truncated'' Steinberg groups $\St^{\leq n}(\Phi, B)$ by the set of generators $x_\alpha(a)$, $\alpha\in \Phi$, $a\in B_d$, $d\leq n$, and relations $\text{R}1_{d}$, $\text{R3}^\angle_{d, e}$ and $\text{R3}^\bot_{d, e}$ for all $d,e \leq n$ and $\text{R2}_{d,e}$ for those $d,e \leq n$ for which $d+e \leq n$.

@@

\begin{lemma} \label{claim3}
 In $\St^{\leq n}(\Phi, B)$ relations $\bot_0(d, e)$, $\angle(d, d)$ and $\angle(e, e)$ imply $\bot(d, e)$ for $d\leq e \leq n$.
\end{lemma}
\begin{proof}
 \begin{align*} 
 x_{\alpha+\beta+\gamma}(abt^{e}) = {}^{x_{-\beta}(t^d)}\!x_{\alpha+\beta+\gamma}(abt^{e}) & \text{ by $\text{R3}^{\bot, 0}_{d, e}$}\\
 = {}^{x_{-\beta}(t^d)}\![x_{\alpha+\beta}(b't^{e-d}), x_\gamma(at^d)] & \text{ by $\text{R2}_{e-d,d}$} \\
 = [x_{\alpha}(b't^{e}) \cdot x_{\alpha+\beta}(b't^{e-d}), x_\gamma(at^d)] & \text{ by $\text{R2}_{e-d,d}$ and $\text{R3}^\angle_{d, d}$ }\\
 = {}^{x_{\alpha}(b't^{e})}\!x_{\alpha+\beta+\gamma}(abt^{e}) [x_{\alpha}(b't^{e}), x_\gamma(at^d)] & \text{ by~\eqref{eq:H1ii} and $\text{R2}_{e-d,d}$} \\ = x_{\alpha+\beta+\gamma}(abt^{e}) [x_{\alpha}(b't^{e}), x_\gamma(at^d)] & \text{ by $\text{R3}^\angle_{e, e}$. \qedhere}
\end{align*}
\end{proof}
\end{comment}
\begin{align}
 \label{eq:H1ii}  [ab, c] = {}^a[b, c] \cdot [a,c];&\\ %= [a,[b,c]]\cdot [b,c] \cdot [a,c];&\\
 \label{eq:H1iii} [a,c]   = 1    \text{ implies } [a, [b,c]] = [[a,b],{}^bc].&
\end{align}

In the case $1\leq d \leq e \leq n$ the assertion follows from the following computation:
\begin{align*} x_{\alpha+\beta+\gamma}(abt^e) = [x_{\alpha+\beta}(at^{e-d}), [x_{-\beta}(t^d), x_{\beta+\gamma}(b)] ] & \text{ by $\text{R2}_{d,0}$, $\text{R2}_{d, e-d}$  } \\ 
 = [[x_{\alpha+\beta}(at^{e-d}), x_{-\beta}(t^d)], {}^{x_{-\beta}(t^d)} x_{\beta+\gamma}(b)] & \text{ by~\eqref{eq:H1iii}, $\text{R3}^\bot_{0,e-d}$} \\ 
 = {}^{x_{-\beta}(t^d)} [x_{\alpha}(at^e), x_{\beta+\gamma}(b)] & \text{ by $\text{R2}_{e-d,d}$, $\text{R3}^\angle_{d,e}$} \\ 
 = {}^{x_{-\beta}(t^d)} [x_{\alpha + \beta}(at^{e-d}), x_{\gamma}(bt^d)] & \text{ by $\text{R2}_{e,0}$, $\text{R2}_{e-d,d}$} \\ 
 = [x_{\alpha}(at^e) \cdot x_{\alpha + \beta}(at^{e-d}), x_{\gamma}(bt^d)] & \text{ by $\text{R2}_{e-d,d}$, $\text{R3}^\angle_{d,d}$} \\ 
 = {}^{x_\alpha(at^e)}x_{\alpha+\beta+\gamma}(abt^e) \cdot [x_\alpha(at^e), x_\gamma(bt^d)] & \text{ by~\eqref{eq:H1ii}, $\text{R2}_{e-d,d}$} \\
 = x_{\alpha+\beta+\gamma}(abt^e) \cdot [x_\alpha(at^e), x_\gamma(bt^d)] & \text{ by $\text{R3}^\angle_{e,e}$}. \end{align*}

In the case $d < 1\leq e \leq n$
\begin{align*} 
 [x_\alpha(\xi), x_{\gamma}(bt^{e})] = [x_\alpha(\xi), [x_{\beta+\gamma}(b't^{e-1}), x_{-\beta}(t)]] & \text{ by~$\text{R2}_{e-1,1}$ } \\
 = [[x_\alpha(\xi), x_{\beta+\gamma}(b't^{e-1})], {}^{x_{\beta+\gamma}(b't^{e-1})}\!x_{-\beta}(t)] & \text{ by \eqref{eq:H1iii} and $\text{R3}^\angle_{d,1}$ } \\
 = {}^{x_{\beta+\gamma}(b't^{e-1})}\![x_{\alpha+\beta+\gamma}(b't^{e-1}\xi), x_{-\beta}(t)] & \text{ by $\text{R2}_{d,e-1}$ and $\text{R3}^\angle_{e-1,e-1}$  } \\
 = 1 & \text{ by $\text{R3}^\bot_{d+e-1,1}$.} \end{align*}

@@

Now the same argument as in Lemma~\ref{welldef} allows to prove that $[s_{\alpha_{l-1}}(a),\,x_{\alpha_l}(b)]=1$ for $b\in R_e$, $e\leq0$, what finishes the proof of R3.2$\!\,_{n+1}$.

\subsection{Computation of the kernel of the map of evaluation at \texorpdfstring{$0$}{0}.}
For a pair $(R, I)$ we denote by $\overline{\St}(\Phi, R, I)$ the image of the map $\mu$ from~\eqref{relSteinberg-def}, in particular, it is a normal subgroup of $\St(\Phi, R)$.
We also denote by $\St(\Phi, I)$ the subgroup of $\St(\Phi, R)$ generated as a group by root unipotents of level $I$.
It is clear that $\overline{\St}(\Phi, R, I)$ contains $\St(\Phi, I)$ and, in fact, is its normal closure.
We also denote by $\overline{H}(\Phi, R, I)$ the subgroup of $\overline{\St}(\Phi, R, I)$ generated by the semisimple root elements $h_\alpha(u)$ and symbols $\{u, v\}$, $u \in (1+I)^*$, $v \in R^*$, $\alpha\in \Phi$.

>>>>

Whenever the coefficient $\xi$ of a Steinberg generator $g = x_\alpha(\xi)$ of $\St(\Phi, B)$ is a homogeneous element of $B$, i.\,e. $\xi \in B_d$ for some $d \in \mathbb{Z}$,
 we call the corresponding generator $g$ homogeneous of degree $d$.
It is not hard to show that $\St(\Phi, B)$ can be presented by the set of all homogeneous Steinberg generators modulo the following set of Steinberg relations
 between them (below $a, a' \in B_d$, $b\in R_e$ and $d,e \in \mathbb{Z}$): 
\begin{align}
&\,\,x_{\alpha}(a)x_{\alpha}(a') =  x_{\alpha}(a+a'),& \tag{R$1_d$} \\
&\,[x_{\alpha}(a),\,x_{\beta}(b)]= x_{\alpha+\beta}(N_{\alpha, \beta} \cdot ab),& \alpha + \beta \in \Phi \tag{R$2_{d,e}$} \\
&[x_{\alpha}(a),\,x_{\beta}(b)]= 1,& \alpha + \beta \not \in \Phi\cup\{0\}.\ \tag{R$3_{d,e}$}
%&[x_{\alpha}(a),\,x_{\beta}(b)]= 1,& \alpha - \beta \in \Phi,\ \tag{R$3^\angle_{d,e}$}
\end{align}
%For such a ring $R$ one can assume in the definition of the Steinberg group that the generators $x_\alpha(a)$ and relations R1--R3 are defined only for homogeneous elements.

For $n\geq 1$ we define ``truncated'' Steinberg groups $\St^{\leq n}(\Phi, B)$ by the set of homogeneous Steinberg generators of degree $\leq n$
modulo relations $\text{R}1_{d}$, $\text{R3}_{d, e}$ for all $d,e \leq n$ and $\text{R2}_{d,e}$ for those $d,e \leq n$ for which $d+e \leq n$.

It will be convenient for us to distinguish between the cases when depending on whether the roots $\alpha, \beta$ appearing in $\text{R3}_{d,e}$ are orthogonal to each other or form a sharp angle.
In the former case we denote such relator by $\text{R3}^\bot_{d,e}$, and in the latter case by $\text{R3}^\angle_{d,e}$.

The following lemma is inspired by the proof of the main result in ???.
Basically, it says that most of the relators $\text{R3}^\bot_{d,e}$ in the presentation of $\St^{\leq n}(\Phi, B)$ are superflous.
\begin{lemma}\label{superfluous-relations}

\end{lemma}
\begin{proof}
We will use the following commutator identities (cf Rehmann-Soule).
\begin{align} \label{eq:H1ii}  [ab, c] = {}^a[b, c] \cdot [a,c];&\\ %= [a,[b,c]]\cdot [b,c] \cdot [a,c];&\\
 \label{eq:H1iii} [a,c]   = 1    \text{ implies } [a, [b,c]] = [[a,b],{}^bc].& \end{align}

In the case $0 < d \leq e \leq n$ the assertion follows from the following computation:
\begin{align*} x_{\alpha+\beta+\gamma}(abt^e) = [x_{\alpha+\beta}(at^{e-d}), [x_{-\beta}(t^d), x_{\beta+\gamma}(b)] ] & \text{ by $\text{R2}_{d,0}$, $\text{R2}_{d, e-d}$  } \\ 
 = [[x_{\alpha+\beta}(at^{e-d}), x_{-\beta}(t^d)], {}^{x_{-\beta}(t^d)} x_{\beta+\gamma}(b)] & \text{ by~\eqref{eq:H1iii}, $\text{R3}^\bot_{0,e-d}$} \\ 
 = {}^{x_{-\beta}(t^d)} [x_{\alpha}(at^e), x_{\beta+\gamma}(b)] & \text{ by $\text{R2}_{e-d,d}$, $\text{R3}^\angle_{d,e}$} \\ 
 = {}^{x_{-\beta}(t^d)} [x_{\alpha + \beta}(at^{e-d}), x_{\gamma}(bt^d)] & \text{ by $\text{R2}_{e,0}$, $\text{R2}_{e-d,d}$} \\ 
 = [x_{\alpha}(at^e) \cdot x_{\alpha + \beta}(at^{e-d}), x_{\gamma}(bt^d)] & \text{ by $\text{R2}_{e-d,d}$, $\text{R3}^\angle_{d,d}$} \\ 
 = {}^{x_\alpha(at^e)}x_{\alpha+\beta+\gamma}(abt^e) \cdot [x_\alpha(at^e), x_\gamma(bt^d)] & \text{ by~\eqref{eq:H1ii}, $\text{R2}_{e-d,d}$} \\
 = x_{\alpha+\beta+\gamma}(abt^e) \cdot [x_\alpha(at^e), x_\gamma(bt^d)] & \text{ by $\text{R3}^\angle_{e,e}$}. \end{align*}

In the case $d \leq 0 \leq e \leq n$
\begin{align*} 
 [x_\alpha(\xi), x_{\gamma}(bt^{e})] = [x_\alpha(\xi), [x_{\beta+\gamma}(b't^{e-1}), x_{-\beta}(t)]] & \text{ by~$\text{R2}_{e-1,1}$ } \\
 = [[x_\alpha(\xi), x_{\beta+\gamma}(b't^{e-1})], {}^{x_{\beta+\gamma}(b't^{e-1})}\!x_{-\beta}(t)] & \text{ by \eqref{eq:H1iii} and $\text{R3}^\angle_{d,1}$ } \\
 = {}^{x_{\beta+\gamma}(b't^{e-1})}\![x_{\alpha+\beta+\gamma}(b't^{e-1}\xi), x_{-\beta}(t)] & \text{ by $\text{R2}_{d,e-1}$ and $\text{R3}^\angle_{e-1,e-1}$  } \\
 = 1 & \text{ by $\text{R3}^\bot_{d+e-1,1}$. \qedhere} \end{align*}
\end{proof}

@@

\begin{lemma} \label{claim3}
 In $\St^{\leq n}(\Phi, B)$ relations $\bot_0(d, e)$, $\angle(d, d)$ and $\angle(e, e)$ imply $\bot(d, e)$ for $d\leq e \leq n$.
\end{lemma}
\begin{proof}
 \begin{align*} 
 x_{\alpha+\beta+\gamma}(abt^{e}) = {}^{x_{-\beta}(t^d)}\!x_{\alpha+\beta+\gamma}(abt^{e}) & \text{ by $\text{R3}^{\bot, 0}_{d, e}$}\\
 = {}^{x_{-\beta}(t^d)}\![x_{\alpha+\beta}(b't^{e-d}), x_\gamma(at^d)] & \text{ by $\text{R2}_{e-d,d}$} \\
 = [x_{\alpha}(b't^{e}) \cdot x_{\alpha+\beta}(b't^{e-d}), x_\gamma(at^d)] & \text{ by $\text{R2}_{e-d,d}$ and $\text{R3}^\angle_{d, d}$ }\\
 = {}^{x_{\alpha}(b't^{e})}\!x_{\alpha+\beta+\gamma}(abt^{e}) [x_{\alpha}(b't^{e}), x_\gamma(at^d)] & \text{ by~\eqref{eq:H1ii} and $\text{R2}_{e-d,d}$} \\ = x_{\alpha+\beta+\gamma}(abt^{e}) [x_{\alpha}(b't^{e}), x_\gamma(at^d)] & \text{ by $\text{R3}^\angle_{e, e}$. \qedhere}
\end{align*}
\end{proof}

@@

Now the same argument as in Lemma~\ref{welldef} allows to prove that $[s_{\alpha_{l-1}}(a),\,x_{\alpha_l}(b)]=1$ for $b\in R_e$, $e\leq0$, what finishes the proof of R3.2$\!\,_{n+1}$.
\end{comment}
\subsection{Computation of the kernel of the map of evaluation at \texorpdfstring{$0$}{0}.}
For a pair $(R, I)$ we denote by $\overline{\St}(\Phi, R, I)$ the image of the map $\mu$ from~\eqref{relSteinberg-def}, in particular, it is a normal subgroup of $\St(\Phi, R)$.
We also denote by $\St(\Phi, I)$ the subgroup of $\St(\Phi, R)$ generated as a group by root unipotents of level $I$.
It is clear that $\overline{\St}(\Phi, R, I)$ contains $\St(\Phi, I)$ and, in fact, is its normal closure.
We also denote by $\overline{H}(\Phi, R, I)$ the subgroup of $\overline{\St}(\Phi, R, I)$ generated by the semisimple root elements $h_\alpha(u)$ and symbols $\{u, v\}$, $u \in (1+I)^*$, $v \in R^*$, $\alpha\in \Phi$.
